---
title: "Journal Club Notes: Probabilistic Weather Forecasting in R"
author: "Cory Cutsail, PhillyRUG"
---

Two R packages introduced: `ensembleMBA` and `ProbForecastGOP`.

`ensebmbleBMA`:

-   Implements geostatistical output perturbation method

-   Use mixture distributions where components $X \iff \text{ensemble members}$

-   $X\sim D_j$

-   $j\in\{\text{temp,precip,wind speed,...}\}$

`ProbForecastGOP`:

-   Geostatistical methods to produce prob. forecasts of entire weather fields based on a single numerical forecast on a spatial grid

-   Package removed from CRAN ... we can talk about the math but we'll skip the code discussion

Loading `ensembleBMA`

```{r echo=F, include=T}
#install.packages('ensembleBMA')
library('ensembleBMA')
# install.packages('ProbForecastGOP')
# library('ProbForecastGOP')
```

Parameter estimation done via EM (fairly simple algo .. we can pause here but the [Wikipedia article](https://en.wikipedia.org/wiki/Expectation–maximization_algorithm) covers it nicely).

Data used for training must be an `ensembleBMA::ensembleData` object. The authors use data from [Berrocal et al, 2007](https://sites.stat.washington.edu/raftery/Research/PDF/Berrocal2007.pdf) to illustrate the construction of an `ensembleData` object.

```{r echo=F,include=T}
data(srft)
members <- c("CMCG", "ETA", "GASP", "GFS",
             "JMA", "NGPS", "TCWB", "UKMO")
srftData <-
  ensembleData(forecasts = srft[,members],
    dates = srft$date,
    observations = srft$obs,
    latitude = srft$lat,
    longitude = srft$lon,
    forecastHour = 48)
```
